---
title: "Final Report"
author: "Student ID 2935413"
date: "2023-12-13"
bibliography: refs.bib
csl: apa.csl
nocite: '@*'
output: 
  bookdown::html_document2:
    toc: true
    toc_float :
      collapsed: false
    number_sections: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse) # for data wrangling
library(kableExtra) # for building tables
library(faux) # for correlation simulation

# ggplot theme
theme_set(theme_minimal(base_size = 15))
```

# Descriptives 

```{r data-load}
# reading the raw data from csv files
sss_orig <- read_csv("data/sss_orig.csv", 
                     show_col_types = FALSE, 
                     na = c("NULL"))
state_orig <- read_csv("data/state_orig.csv", show_col_types = FALSE)
state_q <- read_csv("data/state_q.csv", show_col_types = FALSE)
subjects <- read_csv("data/subjects.csv", show_col_types = FALSE)
trait_orig <- read_csv("data/trait_orig.csv", show_col_types = FALSE)
trait_q <- read_csv("data/trait_q.csv", show_col_types = FALSE)
```

```{r data-processing-state}
# converting state_orig to long format
state_long <- state_orig |>
  pivot_longer(cols = "calm":"pleasant", 
               names_to = "q_id",
               values_to = "response") 

# joining the state_long with the state_q dataset
merged_state <- left_join(state_long, state_q, by = "q_id")
  
# score for state anxiety
response_score <- c("Not at all" = 1,
                    "Somewhat" = 2,
                    "Moderately so" = 3, 
                    "Very much so" = 4)

state_scores <- merged_state |> 
  mutate(unreversed_score = response_score[response], 
         score = ifelse(reversed == TRUE, 
                        5 - unreversed_score, 
                        unreversed_score)) |>
  group_by(user_id) |>
  summarize(state_score = sum(score)) |>
  distinct(user_id, state_score)
```

```{r data-processing-trait}
# converting trait_orig to long format
trait_long <- trait_orig |>
  pivot_longer(cols = "pleasant" : "turmoil", 
               names_to = "q_id", 
               values_to = "response")

# joining the trait_long with the trait_q dataset
merged_trait <- left_join(trait_long, trait_q, by = "q_id")

# score for trait anxiety
trait_scores <- merged_trait |> 
  mutate(unreversed_score = response_score[response], 
         score = ifelse(reversed == TRUE, 
                        5 - unreversed_score, 
                        unreversed_score)) |>
  group_by(user_id) |>
  summarize(trait_score = sum(score)) |>
  distinct(user_id, trait_score)
```

``` {r data-processing-sss}
# converting to long format
sss_long <- sss_orig |>
  pivot_longer(cols = "sss-1":"sss-14", 
               names_to = "question",
               values_to = "score")

# score for sensation seeking
sensation_scores <- sss_long |> 
  mutate(sen_score = ifelse(score == TRUE, 1, 0))|>
  group_by(user_id) |>
  summarize(sensation_score = sum(sen_score)) |>
  distinct(user_id, sensation_score)
```

## Demographics

```{r join-scores}
# participants who started all three questionnaires
final_subjects <- inner_join(state_scores, trait_scores, by = "user_id") |> 
  inner_join(sensation_scores, by = "user_id") |>
  inner_join(subjects, by = "user_id") 

# calculate demographics values
tot_subj <- nrow(subjects)
sss_subj <- nrow(sss_orig)
state_subj <- nrow(state_orig)
trait_subj <- nrow(trait_orig)
included_subj <- nrow(final_subjects)
completed_subj <- final_subjects |> 
  complete.cases()|>
  sum()
```

`r tot_subj` people participated in total, with `r sss_subj` starting the sensation seeking questionnaire, `r state_subj` starting the state anxiety questionnaire, and `r trait_subj` starting the trait anxiety questionnaire. After excluding participants who did not start all 3 questionnaires, there were `r included_subj` people left for analyses. `r completed_subj` of these participants completed all 3 questionnaires.

```{r summary-table}
# making a data table with summary statistics for each group
summary_data <- final_subjects |>
  group_by(group, condition)|>
  summarize(n_state = sum(!is.na(state_score)),
            mean_state = mean(state_score, na.rm = TRUE),
            sd_state = sd(state_score, na.rm = TRUE),
            n_trait = sum(!is.na(trait_score)),
            mean_trait = mean(trait_score, na.rm = TRUE),
            sd_trait = sd(trait_score, na.rm = TRUE),
            n_sensation = sum(!is.na(sensation_score)),
            mean_sensation = mean(sensation_score, na.rm = TRUE),
            sd_sensation = sd(sensation_score, na.rm = TRUE)) 

# creating a summary statistics table
summary_table <- summary_data |>
  kable(booktabs = TRUE, digits = 2, 
        caption = "Summary statistics of state anxiety, trait anxiety, and sensation seeking scores for each group.",
        col.names = c("Group", "Condition", 
                      "N", "Mean", "S.D",
                      "N", "Mean", "S.D",
                      "N", "Mean", "S.D")) |>
  kable_styling(full_width = FALSE) |>
  add_header_above(c(" " = 1, " " = 1,
                     "State Anxiety" = 3,
                     "Trait Anxiety" = 3,
                     "Sensation Seeking" = 3))

summary_table
```
## Score Distributions

```{r state-dist, fig.cap = "Distribution of state anxiety score for each group/ condition combination."}
# concatenating group and condition to make graphs
processed_data <- final_subjects |>
  mutate(grp_condn = paste(group, condition, sep = "/"))

# violin plot for state score
state_plot <- processed_data |>
  ggplot(aes(grp_condn, state_score, fill = grp_condn)) +
  geom_violin(draw_quantiles = 0.5, 
              trim = FALSE, 
              scale = "width", 
              width = 0.8) +
  scale_fill_manual(values = c("darkorchid", "darkorange",
                               "dodgerblue", "firebrick")) +
  labs(x = "Group/Condition Combinations", 
       y = "State Anxiety Score") +
  guides(fill = "none") +
  coord_cartesian(ylim = c(0, 80))

state_plot
```

```{r trait-dist, fig.cap = "Distribution of trait anxiety score for each group/ condition combination."}
# violin plot for trait score
trait_plot <- processed_data |>
  ggplot(aes(grp_condn, trait_score, fill = grp_condn)) +
  geom_violin(draw_quantiles = 0.5, 
              trim = FALSE, 
              scale = "width", 
              width = 0.8) +
  scale_fill_manual(values = c("darkorchid", "darkorange",
                               "dodgerblue", "firebrick")) +
  labs(x = "Group/Condition Combinations", 
       y = "Trait Anxiety Score") +
  guides(fill = "none") +
  coord_cartesian(ylim = c(0, 80))

trait_plot
```

```{r sss-dist, fig.cap = "Distribution of sensation seeking score for each group/ condition combination."}
# violin plot for sensation seeking score
sss_plot <- processed_data |>
  ggplot(aes(grp_condn, sensation_score, fill = grp_condn)) +
  geom_violin(draw_quantiles = 0.5, 
              trim = FALSE, 
              scale = "width", 
              width = 0.8) +
  scale_fill_manual(values = c("darkorchid", "darkorange",
                               "dodgerblue", "firebrick")) +
  labs(x = "Group/Condition Combinations", 
       y = "Sensation Seeking Score") +
  guides(fill = "none") +
  coord_cartesian(ylim = c(0, 20))

sss_plot
```

# Analysis

## Correlation

```{r analysis-corr}
# conducting correlation analysis between state and trait anxiety for A/control
analysis_data <- filter(processed_data, grp_condn == "A/control")
corr <- cor.test(analysis_data$state_score, analysis_data$trait_score)

# extracting correlation results
esti = round(corr$estimate, 2)
df = corr$parameter
p = corr$p.value
prob = ifelse(p < .001, "<", "=")
p = ifelse(p < .001, .001, round(p, 3))
ci = round(corr$conf.int, 2)
direction = ifelse(esti < 0, "negatively", "positively")
```

Figure \@ref(fig:fig-corr) shows the correlation between state and trait anxiety scores for group A/ control condition combination. According to the Pearson's correlation test, state and trait anxiety were `r direction` correlated: $r_{`r df`}$ = `r esti`, p `r prob` `r p`, 95% CI = [`r ci`]. 

```{r fig-corr, fig.cap = "The relationship between state and trait anxiety scores for Group A/ control condition."}
# visualizing correlation using a scatterplot
scatterplot_corr <- ggplot(analysis_data, 
                           aes(state_score, trait_score)) + 
  geom_point(color = "darkorchid") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  labs(x = "State Anxiety Score", 
       y = "Trait Anxiety Score") +
  scale_x_continuous(limits = c(20, 80),
                     breaks = seq(20, 80, by = 10)) +
  scale_y_continuous(limits = c(20, 80),
                     breaks = seq(20, 80, by = 10)) 
scatterplot_corr
```

## Statistical Power

```{r power-sim}
# creating a function to simulate the correlation result
sim_corr_test <- function(n = 100, r = 0,
                          mu = c(0, 20), 
                          sd = c(1, 5),
                          alpha = 0.01) {
  dat <- rnorm_multi(
    n = n, 
    r = r,
    vars = 2,
    varnames = c("x", "y")
  )
  
  ctest = cor.test(dat$x, dat$y)
  sig = ctest$p.value < alpha
  
  return(sig)
}

# number of observations in the A/control combination used for correlation analysis
n_data <- nrow(analysis_data)

# extracting means and sds for state and trait anxiety scores for A/control combination
sim_data <- summary_data |>
  filter(group == "A",
         condition == "control")

mu_state <- pull(sim_data, mean_state) |> round(2)
mu_trait <- pull(sim_data, mean_trait) |> round(2)
sdev_state <- pull(sim_data, sd_state) |> round(2)
sdev_trait <- pull(sim_data, sd_trait) |> round(2)

# assigning value to alpha for reproducibility
alpha = 0.01

# calculating power through iteration
power <- replicate(1e4, sim_corr_test(n = n_data, r = esti,
                                      mu = c(mu_state, mu_trait), 
                                      sd = c(sdev_state, sdev_trait),
                                      alpha = alpha)) |>
  mean() 
```

The statistical power of the correlation test when alpha = `r alpha` is `r power`, which means that the probability of rejecting the null hypothesis (i.e. no correlation exists between state and trait anxiety scores) when it is false is 100%. In other words, the likelihood of successfully detecting a correlation between state and trait anxiety scores when it is actually present is 100%. 

In the power simulation, the parameters used were:

  (1) number of participants = `r n_data`, 
  (2) correlation coefficient = `r esti` as the effect size, 
  (3) alpha = `r alpha`, 
  (4) means of variable x = `r mu_state` and variable y = `r mu_trait`, 
  (5) standard deviations of variable x = `r sdev_state` and variable y = `r sdev_trait`.

# References
